(Re)building cooperation: Effects of a cognitive intervention on cooperative behavior in games

### 2024-04-19 article.Rmd
ready for submission

---
title: "(Re)building cooperation: Effects of a cognitive intervention on cooperative behavior in games"
# author:
#   "Ismail Guennouni^1^, Samuel Dupret^1^, Quentin JM Huys^2^, Maarten Speekenbrink^1^"

@@

bibliography: "bib/cogIntervention.bib"
biblio-style: spphys

output:
  bookdown::pdf_document2:
    toc: false
header-includes:
  - \usepackage{lscape}
---
<!-- \small   -->
<!-- ^1^ *Department of Experimental Psychology, Division of Psychology and Language Sciences, UCL*.   -->
<!-- ^2^ *Division of Psychiatry and Max Planck Centre for Computational Psychiatry and Ageing Research, Queen Square Institute of Neurology, UCL.* -->

@@

## Research Transparency Statement {.unnumbered}

### General Disclosures {.unnumbered}

Conflicts of interest: The author(s) declared that there were no conflicts of interest with respect to the authorship or the publication of this article.
Funding: Author 1 was supported by the UK Engineering and Physical Sciences Research Council under grant EP/S515255/1. Author 3 acknowledges support by the UCLH NIHR BRC. He has received fees and options for consultancies for Aya Technologies and Alto Neuroscience.
Ethics: This research received approval from a local ethics board (ID:21029/001).
Artificial intelligence: No artificial intelligence assisted technologies were used in this research or the creation of this article.
Reproducibility: [The authors are applying for a Computational Reproducibility Badge which will be awarded pending checks by the STAR Team.]

### Study Disclosures {.unnumbered}

Preregistration: The hypotheses and methods were not preregistered.The analysis plan was not preregistered. Materials: All study materials are publicly available (https://github.com/ismailg/coax_psych_science). Data: All primary data are publicly available (https://github.com/ismailg/coax_psych_science). Analysis scripts: All analysis scripts are publicly available (https://github.com/ismailg/coax_psych_science).

<!-- Author Contribution: I. Guennouni, QJM. Huys and M. Speekenbrink designed and developed the study concept. Experiment design was done by S. Dupret and I. Guennouni. Testing and data collection were performed by I. Guennouni. I. Guennouni analysed and interpreted the data under the supervision of QJM. Huys and M. Speekenbrink. M. Speekenbrink and I. Guennouni jointly performed the HMM modelling. All authors jointly wrote and approved the final version of the manuscript for submission. -->

# Introduction

@@

# Method

## Participants

@@

## Design and Procedure

The experiment had a 2 (Condition: Intervention or Control) by 3 (Game: Trust-Game Pre Intervention, Trust-Game Post Intervention, Prisoner's Dilemma Post Intervention) design, with repeated measures on the second factor. Participants were randomly assigned to one of the two levels of the first factor. The games were designed and implemented online using Empirica [@almaatouq2021empirica].

@@

## Tasks and Measures

### Repeated Trust Game

Participants played a 15-round Repeated Trust Game [@joyce_trust_1995] in the trustee role against a computer-programmed investor. On each round the investor is endowed with 20 units and decides how much of that endowment to invest. This investment is tripled and the trustee then decides how to split this tripled amount between them and the investor. If the trustee returns more than one third of the amount, the investor makes a gain. Each player was represented with an icon with the participant always on the left of the screen and the co-player on the right (Figure \ref{fig:HMMPanels}.A). The participants were able to choose the icon that represents them at the start of the experiment. The icon representing the co-player changed at the start of each new game, to simulate a new interaction partner. Participants were not told they were facing computerised co-players.

The strategy of the computerised investor was modelled on behavior of human investors in the Repeated Trust Game (RTG)  over 10-rounds with the same (human) co-player. Full detail on the datasets used in the Supplementary Information. Using this data, we estimated a hidden Markov model (HMM) on investors' behavior with three latent states. Each latent state was associated with a state-conditional distribution over the possible investments from 0 to 20 (Figure \ref{fig:HMMPanels}.C). These distributions reflect "low-trust", "medium-trust", or "high-trust". Over rounds, the investor can move between states, and the probability of these transitions was modelled as a function of their net return (i.e return - investment) in the previous round (see Figure \ref{fig:HMMPanels}.D). In order to instigate a potential breakdown of trust, thereby allowing us to probe efforts to repair trust, the computerised agent was programmed to provide a low investment on round 12 (pre-intervention) or round 13 (post-intervention). On all other rounds, the investor's actions were determined by randomly drawing an investment from the state-conditional distribution, with the state over rounds determined by randomly drawing the next state from the state-transition distribution as determined from the net return on the previous round (disregarding the net return immediately after the pre-programmed low investment rounds). The initial state for the HMM investor in each instance of the game was the "mid-trust" state.

On each round, immediately after being informed of the investment sent, participants in the intervention condition were asked to provide an evaluation of their emotion in terms of valence (from negative to positive) and arousal (from low to high). Participants in the control condition were asked to evaluate the investment in terms of speed (from slow to fast) and magnitude (from low to high). These evaluations were made by clicking on a two-dimensional field with labelled axes as shown in Figure \ref{fig:HMMPanels}.C for the intervention condition.

### Repeated Prisoner's Dilemma

@@

## Intervention

@@

## Procedure

@@

## Statistical analysis

@@

The model was estimated using the `afex` package [@singmann_afex_2022] in R [@R_package]. More complex models with additional random effects could not be estimated reliably, and as such the estimated model can be considered to include the optimal random effects structure [@matuschek2017balancing]. A similar process was used to establish the random effects structures of other linear mixed-effects models used throughout the statistical analyses. For the $F$-tests, we used the Kenward-Roger approximation to the degrees of freedom, as implemented in the R package "afex". We Z-transform the Investment variable as centering is beneficial to interpreting the main effects more easily in the presence of interactions.

@@

\begin{landscape}
```{r HMMPanels, fig.cap= "\\small{A: Screenshot of the repeated Trust Game. The game is played in dyads, with one player assigned the role of the investor and the other player that of the trustee. The investor is endowed with 20 units at the start of each round. They need to decide how much of that endowment they want to invest with the trustee. The investment is then multiplied by a factor of 3 and sent to the trustee who needs to decide how much of the multiplied investment they want to send back to the investor. Shown here is the stage at which the trustee decides how much to send back to the investor. Immediately before making this decision, participant are shown a two dimensional grid (Panel C) where they need to report their reaction to the investment. Those in the intervention condition were asked questions about their emotional state and those in the control condition were asked to report features of the investment unrelated to their emotion. Panels B - D: We construct the artificial investor agent by fitting a three-state hidden Markov model to data of human investors engaged in the 10 round Repeated Trust Game against human trustees. From the fitted HMM, we get the distribution of investments by the artificial investor agent conditional on its latent state as shown in Panel C. The fitted HMM also yields the transition probability of the agent to a state on trial t+1 as a function of the net return (difference between the investment sent and the amount received in return) on trial t as shown in Panel D. Each plot in Panel D represents a different starting latent state on trial t, and each line represents the probability of transitioning to a particular state in trial t+1.}",echo=FALSE, message=FALSE, warning=FALSE, out.width='1\\linewidth', out.height='1\\textheight', fig.align='center'}

@@

# Annotate each image with a label
image1 <- image_annotate(image1, text = "A", location = "+200+10", size = 90, color = "black", gravity = "northwest")
image2 <- image_annotate(image2, text = "B", location = "+0+10", size = 90, color = "black", gravity = "northwest")
image3 <- image_annotate(image3, text = "C", location = "-0+10", size = 90, color = "black", gravity = "northwest")
image4 <- image_annotate(image4, text = "D", location = "+0+10", size = 90, color = "black", gravity = "northwest")

@@

# Behavioral results

@@

mod_returns_pct <- mixed( ret_pct ~ gameNum.f*condition.f*inv_scaled+ (1 + gameNum.f| playerId), avg_ret_df, REML= TRUE, method="KR")
anova(mod_returns_pct)
summary(mod_returns_pct)

@@

```{r modemmeans, include=F, cache=TRUE }
prepost_bycond <- pairs(emmeans::emmeans(mod_returns_pct, c("gameNum.f"), by = "condition.f", pbkrtest.limit = 9400))
cond_effect <- pairs(emmeans::emmeans(mod_returns_pct, c("condition.f"), by="gameNum.f"))
```

<!-- Secondary analyses for effect size and correlation between repeated measures.  -->
```{r, include=F}

library(sjstats)
effectsize::eta_squared(mod_returns_pct)

# Subset the data for pre and post measurements
pre_data <- avg_ret_df$ret_pct[avg_ret_df$gameNum.f == 'pre']
post_data <- avg_ret_df$ret_pct[avg_ret_df$gameNum.f == 'post']

# Calculate the correlation
correlation <- cor(pre_data, post_data, use = "complete.obs")

# Print the result
print(correlation)

@@

```{r, include=F}
# Function to calculate y_position for geom_signif
calculate_y_position <- function(plot) {
  plot_data <- ggplot_build(plot)$data[[1]]
  max_y <- max(plot_data$ymax, na.rm = TRUE)
  y_position <- max_y + (max_y * 0.05) # Adding 5% of the maximum y-value as a margin
  return(y_position)
}
```

@@

Average investments and returns prior to the "defection round" (Figure \ref{fig:TrustGameResults}.A) were within the range of reported investments (40-60% of endowment) and returns (35-50% of total yield) in the literature [@charness_investment_2008; @fiedler_social_2011]. 
Mixed-effects analysis on the percentage returns shows a significant main effect of Condition (intervention vs. control), `r papaja::apa_print(mod_returns_pct)$full_result$condition_f`, with higher percentage returns in the intervention compared to the control condition. Importantly, we also find an interaction between Condition and Game (RTG pre- vs. post-intervention), `r papaja::apa_print(mod_returns_pct)$full_result$gameNum_f_condition_f`. Post-hoc tests show an increase in the percentage returned in the intervention condition, pre - post, `r papaja::apa_print(prepost_bycond)$full_result$Intervention_Pre_post`, but a decrease in the control condition, `r papaja::apa_print(prepost_bycond)$full_result$Control_Pre_post` (see Figure \ref{fig:TrustGameResults}.C). This indicates the intervention was effective in increasing cooperative behavior. There was also a significant main effect of Investment, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct)$full_result$inv_scaled)`,
such that higher investments were associated with higher percentage returns. An Investment by Condition interaction, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct)$full_result$condition_f_inv_scaled)`, indicates that returns were more affected by investments in the control condition. There was also an Investment by Game interaction, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct)$full_result$gameNum_f_inv_scaled)`. 
Finally, we find a three way interaction between Game, Condition and Investment, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct)$full_result$gameNum_f_condition_f_inv_scaled)`, showing that the differentiated effect of the investment on the proportion returned by condition is itself moderated by the Game (pre- vs post intervention).

@@

To explore the HMM investors' behavior across games and conditions, we estimate a linear mixed-effects model of investments sent by the computerised HMM agent with Condition and Game and their interaction as fixed effects, and a similar random effects structure to the returns model. This shows a main effect of Condition, `r papaja::apa_print(mod_invs)$full_result$condition_f`<!--$F(1,317) = 8.7$, $p = .003$-->, and Game, `r papaja::apa_print(mod_invs)$full_result$gameNum_f`<!--$F(1,317) = 8.3$, $p = .004$-->. As can be seen in Figure \ref{fig:TrustGameResults}.D, investment was higher in the intervention compared to the control condition across games, and higher in the second game compared to first across conditions.

<!-- Linear fixed effects model for returns: ALL ROUNDS BEFORE DEFECTION TRIAL -->
```{r, include=FALSE, cache=TRUE}

@@

mod_returns_pre <- mixed( ret_pct ~ gameNum.f*condition.f*inv_scaled+ (1 + gameNum.f| playerId), pre_int_data, REML= TRUE, method="KR")
anova(mod_returns_pre)
summary(mod_returns_pre)
saveRDS(mod_returns_pre, "data/mod_returns_pre")

pairs_pre_int <- pairs(emmeans::emmeans(mod_returns_pre, c("gameNum.f"), by = "condition.f", pbkrtest.limit = 9400, reverse=TRUE))
saveRDS(pairs_pre_int, "data/pairs_pre_int")

@@

@@

mod_postcoax_trust <- mixed(ret_pct ~ gameNum.f*condition.f*inv_scaled  + ( 1 + gameNum.f|playerId), post_coax_data)
anova(mod_postcoax_trust)

postcoax_bygame <- emmeans::emmeans(mod_postcoax_trust, ~gameNum.f, pbkrtest.limit = 9400)
postcoax_bycond <- emmeans::emmeans(mod_postcoax_trust, ~condition.f, pbkrtest.limit = 9400)

postcoax_interaction <- emmeans::emmeans(mod_postcoax_trust, c("gameNum.f"), by = "condition.f", pbkrtest.limit = 9400)
postcoax_interaction

pairs(emmeans::emmeans(mod_postcoax_trust, c("gameNum.f"), by = "condition.f"))
```

```{r, include=F}
saveRDS(summary(mod_postcoax_trust), "data/mod_returns_post")
```

@@

## Emotion self-reports

@@

To assess the impact of the intervention on the participant's emotional reactions, we used linear mixed-effects models (one for valence, and one for arousal) with fixed effects for Game (pre vs. post intervention) and Investment, as well as interaction between Investment and Game, with participant-wide random intercepts and random slopes for Game. This showed that higher investments were associated with more positive emotions, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_emo_x)$full_result$scaleinvestment)`, 
and higher arousal, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_emo_y)$full_result$scaleinvestment)`. 
In addition, the positiveness of emotion declined between the two games,
`r papaja::apa_print(mod_emo_x)$full_result$gameNum_f`
, as did arousal,
`r papaja::apa_print(mod_emo_y)$full_result$gameNum_f`. 
There was no indication that the effect of the investment on either aspect of emotion was affected by the intervention, as there was no interaction between Investment and Game on valence, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_emo_x)$full_result$gameNum_f_scaleinvestment)`, or arousal, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_emo_y)$full_result$gameNum_f_scaleinvestment)`. 
This indicates that participants in the intervention condition returned higher amounts post-intervention, despite their emotional reaction to investments remaining largely the same (Figure \ref{fig:TrustGameResults}.B).

## Evaluation of the investor

@@

## Transfer to the Repeated Prisoner's Dilemma game

@@

\newpage
\begin{landscape}
```{r TrustGameResults, fig.cap="\\small{Panel A: Averages and standard errors of the trustee's return as a percentage of the multiplied investment received by Condition, Phase, and game round. We note a different reaction to the pre-programmed one-off low investment (vertical black lines) between the two conditions: Whilst there is a dip in returns pre-manipulation for both conditions,  post manipulation we see higher returns in the intervention condition compared to the dip in returns seen in the control condition in the right panel. B:  Self-reported emotion valence and arousal as well as investment z-scores for each round of the Repeated Trust Game averaged across participants in the intervention condition only. The participants' emotion reaction measured during the investor's pre-programmed one-off low investment round (vertical grey lines) were similar before (round 12) and after (round 13) the intervention. The bottom panels show marginal means and distributions of either investments or percentage returns across participants by Game and Condition. Panel C shows that participants in the Intervention condition returned higher proportions of the multiplied investment received in the second game compared to the first game over all ronds, whilst those in the Control condition sent back lower returns. Investments by the HMM agent (Panel D) were higher in the second game compared to the first game across conditions.}", echo=FALSE, message=FALSE, warning=FALSE, out.width='1\\linewidth', out.height='1\\textheight', fig.align='center'}
library(magick)

@@

image5 <- image_annotate(image5, text = "A", location = "+20+20", size = 70, color = "black", gravity = "northwest")
image6 <- image_annotate(image6, text = "B", location = "+20+0", size = 90, color = "black", gravity = "northwest")
violin_ret<- image_annotate(violin_ret, text = "C", location = "+0+20", size = 70, color = "black", gravity = "northwest")
violin_inv  <- image_annotate(violin_inv , text = "D", location = "+0+0", size = 70, color = "black", gravity = "northwest")

@@

# HMM analysis of participant returns

@@

Focusing on models with 6 latent states, likelihood ratio tests showed that the HMM-full model fits significantly better than HMM-ctrl ($\chi^2(60) = `r round(-2*as.numeric(logLik(bestCtrl[[6]])) - (-2*as.numeric(logLik(bestFull[[6]]))) , 2)`$, $p < .001$), HMM-coax ($\chi^2(60) = `r round(-2*as.numeric(logLik(bestCoax[[6]])) - (-2*as.numeric(logLik(bestFull[[6]]))) , 2)`$, $p < .001$) and HMM-prepost ($\chi^2(60) = `r round(-2*as.numeric(logLik(bestPrePost[[6]])) - (-2*as.numeric(logLik(bestFull[[6]]))) , 2)`$, $p < .001$). As such, there is evidence that participants reacted differently to the investments in the four combinations of Condition and Game. The estimated state-dependent policy of trustee actions, according to the HMM-full model, is depicted in Figure \ref{fig:trusteeHMM}.A.

>>>>

---
title: "(Re)building cooperation: Effects of a cognitive intervention on cooperative behavior in games"
author:
  "Ismail Guennouni^1^*, Samuel Dupret^1^, Quentin JM Huys^2^, Maarten Speekenbrink^1^"

@@

output:
  bookdown::pdf_document2:
    toc: false
header-includes:
  - \usepackage{lscape}
  # - \usepackage{natbib}

bibliography: "bib/cogIntervention.bib"
csl: nature-no-superscript.csl

---
\small
^1^ *Department of Experimental Psychology, Division of Psychology and Language Sciences, UCL*

^2^ *Division of Psychiatry and Max Planck Centre for Computational Psychiatry and Ageing Research, Queen Square Institute of Neurology, UCL.*

^\*^ *Email: ismail.guennouni@zi-mannheim.de*

@@

# Introduction {.unnumbered}

@@

# Methods {.unnumbered}

## Participants  {.unnumbered}

@@

## Design and Procedure {.unnumbered}

The experiment had a 2 (Condition: Intervention or Control) by 3 (Game: Trust-Game Pre Intervention, Trust-Game Post Intervention, Prisoner's Dilemma Post Intervention) design, with repeated measures on the second factor. Participants were randomly assigned to one of the two levels of the first factor. The games were designed and implemented online using Empirica [@almaatouq2021empirica]. This research received approval from a UCL ethics board (ID:21029/001) and the experiment was performed in accordance with the ethics board guidelines and regulations.

@@

## Tasks and Measures {.unnumbered}

### Repeated Trust Game {.unnumbered}

Participants played a 15-round Repeated Trust Game [@joyce_trust_1995] in the trustee role against a computer-programmed investor. On each round the investor is endowed with 20 units and decides how much of that endowment to invest. This investment is tripled and the trustee then decides how to split this tripled amount between them and the investor. If the trustee returns more than one third of the amount, the investor makes a gain. Each player was represented with an icon with the participant always on the left of the screen and the co-player on the right (Figure \ref{fig:HMMPanels}.a). The participants were able to choose the icon that represents them at the start of the experiment. The icon representing the co-player changed at the start of each new game, to simulate a new interaction partner. Participants were not told they were facing computerised co-players.

The strategy of the computerised investor was modelled on behavior of human investors in the Repeated Trust Game (RTG)  over 10-rounds with the same (human) co-player. Full detail on the datasets used in the Supplementary Information. Using this data, we estimated a hidden Markov model (HMM) on investors' behavior with three latent states. Each latent state was associated with a state-conditional distribution over the possible investments from 0 to 20 (Figure \ref{fig:HMMPanels}.c). These distributions reflect "low-trust", "medium-trust", or "high-trust". Over rounds, the investor can move between states, and the probability of these transitions was modelled as a function of their net return (i.e return - investment) in the previous round (see Figure \ref{fig:HMMPanels}.d). In order to instigate a potential breakdown of trust, thereby allowing us to probe efforts to repair trust, the computerised agent was programmed to provide a low investment on round 12 (pre-intervention) or round 13 (post-intervention). On all other rounds, the investor's actions were determined by randomly drawing an investment from the state-conditional distribution, with the state over rounds determined by randomly drawing the next state from the state-transition distribution as determined from the net return on the previous round (disregarding the net return immediately after the pre-programmed low investment rounds). The initial state for the HMM investor in each instance of the game was the "mid-trust" state.

On each round, immediately after being informed of the investment sent, participants in the intervention condition were asked to provide an evaluation of their emotion in terms of valence (from negative to positive) and arousal (from low to high). Participants in the control condition were asked to evaluate the investment in terms of speed (from slow to fast) and magnitude (from low to high). These evaluations were made by clicking on a two-dimensional field with labelled axes as shown in Figure \ref{fig:HMMPanels}.c for the intervention condition.

### Repeated Prisoner's Dilemma {.unnumbered}

@@

## Intervention {.unnumbered}

@@

## Procedure {.unnumbered}

@@

## Statistical analysis {.unnumbered}

@@

The model was estimated using the `afex` package [@singmann_afex_2022] in R [@R_package]. More complex models with additional random effects could not be estimated reliably, and as such the estimated model can be considered to include the optimal random effects structure [@matuschek2017balancing]. A similar process was used to establish the random effects structures of other linear mixed-effects models used throughout the statistical analyses. There is no agreed upon way to calculate effect sizes for mixed effects models, as such we will report instead on testing differences in marginal means. For the $F$-tests, we used the Kenward-Roger approximation to the degrees of freedom, as implemented in the R package "afex". Unless otherwise stated, the default adjustment method for multiple comparison is Tukey's HSD. We Z-transform the Investment variable as centering is beneficial to interpreting the main effects more easily in the presence of interactions.

@@

\begin{landscape}
```{r HMMPanels, fig.cap= "\\small{a: Screenshot of the repeated Trust Game. The game is played in dyads, with one player assigned the role of the investor and the other player that of the trustee. The investor is endowed with 20 units at the start of each round. They need to decide how much of that endowment they want to invest with the trustee. The investment is then multiplied by a factor of 3 and sent to the trustee who needs to decide how much of the multiplied investment they want to send back to the investor. Shown here is the stage at which the trustee decides how much to send back to the investor. Immediately before making this decision, participant are shown a two dimensional grid (Panel c) where they need to report their reaction to the investment. Those in the intervention condition were asked questions about their emotional state and those in the control condition were asked to report features of the investment unrelated to their emotion. Panels b - d: We construct the artificial investor agent by fitting a three-state hidden Markov model to data of human investors engaged in the 10 round Repeated Trust Game against human trustees. From the fitted HMM, we get the distribution of investments by the artificial investor agent conditional on its latent state as shown in Panel c. The fitted HMM also yields the transition probability of the agent to a state on trial t+1 as a function of the net return (difference between the investment sent and the amount received in return) on trial t as shown in Panel d. Each plot in Panel d represents a different starting latent state on trial t, and each line represents the probability of transitioning to a particular state in trial t+1.}",echo=FALSE, message=FALSE, warning=FALSE, out.width='1\\linewidth', out.height='1\\textheight', fig.align='center'}

@@

# Annotate each image with a label
image1 <- image_annotate(image1, text = "a", location = "+200+10", size = 90, color = "black", gravity = "northwest")
image2 <- image_annotate(image2, text = "b", location = "+0+10", size = 90, color = "black", gravity = "northwest")
image3 <- image_annotate(image3, text = "c", location = "-0+10", size = 90, color = "black", gravity = "northwest")
image4 <- image_annotate(image4, text = "d", location = "+0+10", size = 90, color = "black", gravity = "northwest")

@@

# Behavioral results {.unnumbered}

@@

# mod_returns_pct <- mixed( ret_pct ~ gameNum.f*condition.f*inv_scaled+ (1 + gameNum.f| playerId), avg_ret_df, REML= TRUE, method="KR")
# save(mod_returns_pct, file = "data/mod_returns_pct.RData")

@@

load("data/mod_returns_pct.RData")
anova(mod_returns_pct)
summary(mod_returns_pct)
library(sjstats)
effectsize::eta_squared(mod_returns_pct)

@@

```{r modemmeans, include=F, cache=TRUE }
# prepost_bycond <- pairs(emmeans::emmeans(mod_returns_pct, c("gameNum.f"), by = "condition.f", pbkrtest.limit = 9400))
# save(prepost_bycond, file="data/prepost_bycond.RData")
#cond_effect <- pairs(emmeans::emmeans(mod_returns_pct, c("condition.f"), by="gameNum.f"))
```

```{r}
load("data/prepost_bycond.RData")
```

<!-- VIOLIN PLOTS OF ALL RETURNS MODEL -->
```{r, include=F}
# Function to calculate y_position for geom_signif
calculate_y_position <- function(plot) {
  plot_data <- ggplot_build(plot)$data[[1]]
  max_y <- max(plot_data$ymax, na.rm = TRUE)
  y_position <- max_y + (max_y * 0.05) # Adding 5% of the maximum y-value as a margin
  return(y_position)
}
```

@@

Average investments and returns prior to the "defection round" (Figure \ref{fig:TrustGameResults}.a) were within the range of reported investments (40-60% of endowment) and returns (35-50% of total yield) in the literature [@charness_investment_2008; @fiedler_social_2011]. 
Mixed-effects analysis on the percentage returns shows a significant main effect of Condition (intervention vs. control), `r papaja::apa_print(mod_returns_pct)$full_result$condition_f`, 
<!-- with an effect size (partial eta-squared) of $\eta_p^2$ = `r round(effectsize::eta_squared(mod_returns_pct)$Eta2_partial[2],2)`, -->
due to overall higher percentage returns in the intervention compared to the control condition.
Importantly, we also find an interaction between Condition and Game (RTG pre- vs. post-intervention), `r papaja::apa_print(mod_returns_pct)$full_result$gameNum_f_condition_f`. 
<!-- with an effect size of $\eta_p^2$ = `r round(effectsize::eta_squared(mod_returns_pct)$Eta2_partial[4],2)` -->
Post-hoc tests show an increase in the percentage returned in the intervention condition, pre - post, `r papaja::apa_print(prepost_bycond)$full_result$Intervention_Pre_post`, but a decrease in the control condition, `r papaja::apa_print(prepost_bycond)$full_result$Control_Pre_post` (see Figure \ref{fig:TrustGameResults}.c). This indicates the intervention was effective in increasing cooperative behavior.

There was also a significant main effect of Investment, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct)$full_result$inv_scaled)`, 
<!-- $\eta_p^2$ = `r round(effectsize::eta_squared(mod_returns_pct)$Eta2_partial[3],2)`, -->
such that higher investments were associated with higher percentage returns. An Investment by Condition interaction, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct)$full_result$condition_f_inv_scaled)`, indicates that returns were more affected by investments in the control condition. There was also an Investment by Game interaction, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct)$full_result$gameNum_f_inv_scaled)`. 
Finally, we find a three way interaction between Game, Condition and Investment, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct)$full_result$gameNum_f_condition_f_inv_scaled)`, showing that the differentiated effect of the investment on the proportion returned by condition is itself moderated by the Game (pre- vs post intervention).

@@

To explore the HMM investors' behavior across games and conditions, we estimate a linear mixed-effects model of investments sent by the computerised HMM agent with Condition and Game and their interaction as fixed effects, and a similar random effects structure to the returns model. This shows a main effect of Condition, `r papaja::apa_print(mod_invs)$full_result$condition_f`<!--$F(1,317) = 8.7$, $p = .003$-->, and Game, `r papaja::apa_print(mod_invs)$full_result$gameNum_f`<!--$F(1,317) = 8.3$, $p = .004$-->. As can be seen in Figure \ref{fig:TrustGameResults}.d, investment was higher in the intervention compared to the control condition across games, and higher in the second game compared to first across conditions.

<!-- Linear fixed effects model for returns: ROUNDS BEFORE DEFECTION TRIAL -->
```{r preDefMod, include=FALSE, cache=TRUE}

@@

# mod_returns_pre <- mixed( ret_pct ~ gameNum.f*condition.f*inv_scaled+ (1 + gameNum.f| playerId), pre_int_data, REML= TRUE, method="KR")
# anova(mod_returns_pre)
# summary(mod_returns_pre)
# save(mod_returns_pre, file="data/mod_returns_pre.RData")
# 
# pairs_pre_int <- pairs(emmeans::emmeans(mod_returns_pre, c("gameNum.f"), by = "condition.f", pbkrtest.limit = 9400, reverse=TRUE))
# save(pairs_pre_int, file="data/pairs_pre_int.RData")

@@

```{r loadPreDefMod, include=F}
load("data/mod_returns_pre.RData")
load("data/pairs_pre_int.RData")
```

@@

# mod_postcoax_trust <- mixed(ret_pct ~ gameNum.f*condition.f*inv_scaled  + ( 1 + gameNum.f|playerId), post_coax_data)
# anova(mod_postcoax_trust)
# save(mod_postcoax_trust, file="data/mod_postcoax_trust.RData")
# 
# 
# postcoax_interaction <- emmeans::emmeans(mod_postcoax_trust, c("gameNum.f"), by = "condition.f", pbkrtest.limit = 9400)
# save(postcoax_interaction, file="data/postcoax_interaction.RData")

```{r, include=F}
load("data/mod_postcoax_trust.RData")
load("data/postcoax_interaction.RData")
```

@@

## Emotion self-reports {.unnumbered}

@@

To assess the impact of the intervention on the participant's emotional reactions, we used linear mixed-effects models (one for valence, and one for arousal) with fixed effects for Game (pre vs. post intervention) and Investment, as well as interaction between Investment and Game, with participant-wide random intercepts and random slopes for Game. This showed that higher investments were associated with more positive emotions, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_emo_x)$full_result$scaleinvestment)`, 
and higher arousal, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_emo_y)$full_result$scaleinvestment)`. 
In addition, the positiveness of emotion declined between the two games,
`r papaja::apa_print(mod_emo_x)$full_result$gameNum_f`
, as did arousal,
`r papaja::apa_print(mod_emo_y)$full_result$gameNum_f`. 
There was no indication that the effect of the investment on either aspect of emotion was affected by the intervention, as there was no interaction between Investment and Game on valence, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_emo_x)$full_result$gameNum_f_scaleinvestment)`, or arousal, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_emo_y)$full_result$gameNum_f_scaleinvestment)`. 
This indicates that participants in the intervention condition returned higher amounts post-intervention, despite their emotional reaction to investments remaining largely the same (Figure \ref{fig:TrustGameResults}.b).

## Evaluation of the investor {.unnumbered}

@@

## Transfer to the Repeated Prisoner's Dilemma game {.unnumbered}

@@

\newpage
\begin{landscape}
```{r TrustGameResults, fig.cap="\\small{Panel a: Averages and standard errors of the trustee's return as a percentage of the multiplied investment received by Condition, Phase, and game round. We note a different reaction to the pre-programmed one-off low investment (vertical black lines) between the two conditions: Whilst there is a dip in returns pre-manipulation for both conditions,  post manipulation we see higher returns in the intervention condition compared to the dip in returns seen in the control condition in the right panel. b:  Self-reported emotion valence and arousal as well as investment z-scores for each round of the Repeated Trust Game averaged across participants in the intervention condition only. The participants' emotion reaction measured during the investor's pre-programmed one-off low investment round (vertical grey lines) were similar before (round 12) and after (round 13) the intervention. The bottom panels show marginal means and distributions of either investments or percentage returns across participants by Game and Condition. Panel c shows that participants in the Intervention condition returned higher proportions of the multiplied investment received in the second game compared to the first game over all ronds, whilst those in the Control condition sent back lower returns. Investments by the HMM agent (Panel d) were higher in the second game compared to the first game across conditions.}", echo=FALSE, message=FALSE, warning=FALSE, out.width='1\\linewidth', out.height='1\\textheight', fig.align='center'}
library(magick)

@@

image5 <- image_annotate(image5, text = "a", location = "+20+20", size = 70, color = "black", gravity = "northwest")
image6 <- image_annotate(image6, text = "b", location = "+20+0", size = 90, color = "black", gravity = "northwest")
violin_ret<- image_annotate(violin_ret, text = "c", location = "+0+20", size = 70, color = "black", gravity = "northwest")
violin_inv  <- image_annotate(violin_inv , text = "d", location = "+0+0", size = 70, color = "black", gravity = "northwest")

@@

# HMM analysis of participant returns {.unnumbered}

@@

Focusing on models with 6 latent states, likelihood ratio tests showed that the HMM-full model fits significantly better than HMM-ctrl ($\chi^2(60) = `r round(-2*as.numeric(logLik(bestCtrl[[6]])) - (-2*as.numeric(logLik(bestFull[[6]]))) , 2)`$, $p < .001$), HMM-coax ($\chi^2(60) = `r round(-2*as.numeric(logLik(bestCoax[[6]])) - (-2*as.numeric(logLik(bestFull[[6]]))) , 2)`$, $p < .001$) and HMM-prepost ($\chi^2(60) = `r round(-2*as.numeric(logLik(bestPrePost[[6]])) - (-2*as.numeric(logLik(bestFull[[6]]))) , 2)`$, $p < .001$). As such, there is evidence that participants reacted differently to the investments in the four combinations of Condition and Game. The estimated state-dependent policy of trustee actions, according to the HMM-full model, is depicted in Figure \ref{fig:trusteeHMM}.a.

```{r trusteeHMM, fig.cap="\\small{Panel a: the distribution of participants' percentage return for each of the latent states in the 6 state HMM-full model shows distinct policies centered around different return levels. The latent states are ordered by the mean of the discretized Gaussian representing the policy in that state, so higher numbered states can be considered more pro-social. b: This figure characterises behavioral differences between conditions as the result of participants being in more pro-social (higher return) states in the intervention condition compared to the control condition both pre- and post-defection. The distribution of posterior trustee states post-manipulation by condition for all rounds, as estimated by the most likely posterior state in the best fitting HMM model (HMM-full) using a local decoding procedure is visibly more weighted towards higher return states. As in Panel a, states here are represented in increasing degrees of average percentage returns from the lowest (state 1) to the highest (state 6) return state.}", echo=FALSE, message=FALSE, warning=FALSE, out.width='0.95\\linewidth', out.height='0.80\\textheight', fig.align='center'}
library(magick)
